### 2013-09-19 inst/doc/vignette.Rmd
added accessor methods for eml_read, see README
updating approach based on data.set class

Consider some dataset as an R `data.frame`.

```{r}
dat = data.frame(river = c("SAC", "SAC", "AM"),
                 spp   = c("king", "king", "ccho"),
                 stg   = c("smolt", "parr", "smolt"),
                 ct    =  c(293L, 410L, 210L))
```

Provide a list giving the  __column header__ used in the data.frame, followed by a plain text definitions for the column, followed by a character vector giving the definition for the units:

__column descriptions__: These are usually just plain text definitions, though a URI to a semantic definition can be particularly powerful. See "Advanced Use" for details on adding richer information, such as the method used to collect the data or set the geographic, taxanomic, or temporal coverage of an individual column.

__column units__:   For factors, this is a definition of the levels involved.  For numeric data, specify the units from [this list](http://knb.ecoinformatics.org/software/eml/eml-2.1.1/eml-unitTypeDefinitions.html#StandardUnitDictionary).  For dates, specify the format, (e.g. YYYY or MM-DD-YY). For character strings, a definition of the kind of string can be given, (e.g. species scientific name), otherwise the column description will be used.

```{r}
metadata <- 
  list("river" = list("river",
                      "River site used for collection",
                      c(SAC = "The Sacramento River", 
                        AM = "The American River")),
       "spp" = list("spp",
                    "Species common name", 
                    c(king = "King Salmon", 
                      ccho = "Coho Salmon")),
       "stg" = list("stg",
                    "Life Stage", 
                    c(parr = "third life stage", 
                      smolt = "fourth life stage")),
       "ct"  = list("ct",
                    "count", 
                    "number"))

@@

*for convenience, this metadata list can instead be constructed with the help of the reml wizard*.  Just use `metadata <- metadata_wizard(dat)` to begin. While this may be helpful starting out, regular users will find it faster to define the columns and units directly in the format above.

```{r}
eml_write(dat, metadata, title = "reml example",  
          description = "An example, intended for
                              illustrative purposes only.",
          creator = "Carl Boettiger <cboettig@gmail.com>",
          file = "reml_example.xml")
```

@@

>>>>

`reml` now extends R's `data.frame` class by introducing the `data.set` class which includes additional metadata required by EML.  A `data.set` can be created much like a `data.frame` by specifying additional arguments

```{r}
dat = data.set(river = c("SAC",  "SAC",   "AM"),
               spp   = c("king",  "king", "ccho"),
               stg   = c("smolt", "parr", "smolt"),
               ct    = c(293L,    410L,    210L),
               col.defs = c("River site used for collection",
                            "Species common name",
                            "Life Stage", 
                            "count of live fish in traps"),
               unit.defs = list(c(SAC = "The Sacramento River", 
                                  AM = "The American River"),
                                c(king = "King Salmon", 
                                  ccho = "Coho Salmon"),
                                c(parr = "third life stage", 
                                  smolt = "fourth life stage"),
                                "number"))

@@

- `col.defs`: These are usually just plain text definitions, though a URI to a semantic definition can be particularly powerful. See "Advanced Use" for details on adding richer information, such as the method used to collect the data or set the geographic, taxanomic, or temporal coverage of an individual column.

- `unit.defs`:   For factors, this is a definition of the levels involved.  For numeric data, specify the units from [this list](http://knb.ecoinformatics.org/software/eml/eml-2.1.1/eml-unitTypeDefinitions.html#StandardUnitDictionary).  For dates, specify the format, (e.g. YYYY or MM-DD-YY). For character strings, a definition of the kind of string can be given, (e.g. species scientific name), otherwise the column description will be used.

Alternatively, annotations can be added to an existing data frame, `data.set(my.data.frame, col.defs = ..., unit.defs = ...)`.

```{r}
eml_write(dat, 
          title = "reml example",  
          description = "An example, intended for
                              illustrative purposes only.",
          creator = "Carl Boettiger <cboettig@gmail.com>",
          file = "reml_example.xml")
```

*for convenience, `dat` could simply be a `data.frame` and `reml` will launch it's metadata wizard to assist in constructing the metadata based on the data.frame provided*.  While this may be helpful starting out, regular users will find it faster to define the columns and units directly in the format above.

@@

```{r}
obj <- eml_read("reml_example.xml")
obj 
```

We can use various accessor functions to return the data and metadata elements in native R formats.

```{r}
dataTable(obj)
```

```{r}
metadata <- attributeList(obj)
```

```{r}
contact(obj)
```

Note that the contact has been coerced into R's built-in 'person' object:

```{r}
class(contact(obj))
```

```{r}
citationInfo(obj)
```
